A semantics of core AUTOSAR

### 2016-08-31 paper/paper.tex
Added RAWFP ack. + hidden links
Polishing.

\documentclass[10pt,conference]{IEEEtran}
\usepackage{hyperref}
%\documentclass[twocolumn]{article}
\usepackage{verbatim}
\usepackage[utf8x]{inputenc}
\usepackage[a4paper,margin=1.4cm]{geometry}
\usepackage[fleqn]{amsmath}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage[makeroom]{cancel}
\usepackage{graphicx}
\usepackage{authblk}

@@

An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs) \cite{AR:SWC}. Links between SWCs are called \textbf{connectors}, which attach to \textbf{ports} exposed by the SWCs. A port is either \textbf{required} or \textbf{provided}, and then classified according to what \textbf{interface} it exposes: \textbf{sender-receiver}, \textbf{client-server}, \textbf{trigger} and \textbf{mode-switch} are common interface types. A sender-receiver interface is an aggregation of one or more \textbf{data-elements}, with each element characterized by the type of data it carries and optional further communication specification details. A client-server interface likewise consists of one or more \textbf{operations}, that each give the signature of a procedure call and the possible errors that may result.

@@

%
\[
\begin{array}{rcll}
  i & \in & \text{Component names} \\
  r & \in & \text{Runnable names} \\
  e & \in & \text{Sender-receiver port element names} \\
  o & \in & \text{Client-server port operation names} \\
  s & \in & \text{Inter-runnable variable names} \\
  x & \in & \text{Exclusive area names} \\
\\
  a,b & \in & \text{Addresses} \\
  a,b & ::= & i.r  \sep  i.e   \sep i.o   \sep   i.s   \sep   i.x   \\
\end{array}
\]
%
AUTOSAR elements and operations are actually only unique within the {\em ports} they belong to, so technically we ought to address them by reference to an intended {\em port name} as well, as in $i.p.e$ or $i.p.o$ for some port $p$. However, port names play no role of their own in our study, so we choose to leave them implied by the element and operation names we reference. Moreover, component names \V{I} can also be considered to have a nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc). We can fully ignore this detail here, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.

@@

An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component \cite[ch.~4.3.3.1]{AR:RTE}. Axioms (5) and (7) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$. In both cases the embedded continuation $k$ is applied to the command result. The corresponding \emph{hear} transitions of an irv appear as rules (6) and (8). Thanks to inference rule (1), a runnable instance and an irv can now engage in a parallel transition, for example expressing the writing of shared data.
$$
\begin{array}{c}
  \Tpar {\Trinst {i.r} c \Xs {\TrteIrvWrite s v k}}
        {\Tirv {i.s} \_} \vspace{0.6ex} \\
  \red {\say {i.s} {\Tirvw v}} \\
  \Tpar {\Trinst {i.r} c \Xs {\Tap k \Tvoid}}
        {\Tirv {i.s} v}
\end{array}
$$
The corresponding read transition looks as follows:
$$
\begin{array}{c}
  \Tpar {\Trinst {i.r} c \Xs {\TrteIrvRead s k}}
        {\Tirv {i.s} v} \vspace{0.6ex} \\
  \red {\say {i.s} {\Tirvr v}} \\
  \Tpar {\Trinst {i.r} c \Xs {\Tap k v}}
        {\Tirv {i.s} v}
\end{array}
$$
Section~\ref{sect:Ignore} will show how examples like these can be extended to the case when both participating processes are embedded in arbitrarily large parallel process compositions.

@@

>>>>

\documentclass[10pt,conference]{IEEEtran}
\usepackage[hidelinks]{hyperref}
%\documentclass[twocolumn]{article}
\usepackage{verbatim}
\usepackage[utf8x]{inputenc}
\usepackage[a4paper,margin=1.4cm]{geometry}
\usepackage[fleqn]{amsmath}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage[makeroom]{cancel}
\usepackage{graphicx}
\usepackage{authblk}

@@

An AUTOSAR model is primarily a structure of interconnected \textbf{software components} (SWCs) \cite{AR:SWC}. Links between SWCs are called \textbf{connectors}, which attach to \textbf{ports} exposed by the SWCs. A port is either \textbf{required} or \textbf{provided}, and then classified according to what \textbf{interface} it exposes: \textbf{sender-receiver}, \textbf{client-server}, \textbf{trigger} and \textbf{mode-switch} are common interface types. A sender-receiver interface is an aggregation of one or more \textbf{data-elements}, with each element characterized by the type of data it carries and optional further communication specification details. A client-server interface likewise consists of one or more \textbf{operations}, that each gives the signature of a procedure call and the possible errors that may result.

@@

%
\[
\begin{array}{rcll}
  i & \in & \text{Component names} \\
  r & \in & \text{Runnable names} \\
  e & \in & \text{Sender-receiver port element names} \\
  o & \in & \text{Client-server port operation names} \\
  s & \in & \text{Inter-runnable variable names} \\
  x & \in & \text{Exclusive area names} \\
\\
  a,b & \in & \text{Addresses} \\
  a,b & ::= & i.r  \sep  i.e   \sep i.o   \sep   i.s   \sep   i.x   \\
\end{array}
\]
%
AUTOSAR elements and operations are actually only unique within the {\em ports} they belong to, so technically we ought to address them by reference to an intended {\em port name} as well, as in $i.p.e$ or $i.p.o$ for some port $p$. However, port names play no role of their own in our study, so we choose to leave them implied by the element and operation names we reference. Moreover, component names \V{I} can also be considered to have a nested internal structure (component \V{I_1} within component \V{I_2} within component \V{I_3}, etc). We can ignore this detail too, though, as the hierarchical layout of components in an AUTOSAR system has no run-time implications \cite[ch.~3.3.1]{AR:SWC}.

@@

An inter-runnable variable (irv) represents state that is local to a particular component and accessible by all runnables of that component \cite[ch.~4.3.3.1]{AR:RTE}. Axioms (5) and (7) show the \emph{say} transitions taken by a runnable instance wishing to read and write an irv $s$. In both cases the embedded continuation $k$ is applied to the command result. The corresponding \emph{hear} transitions of an irv appear as axioms (6) and (8). Thanks to inference rule (1), a runnable instance and an irv can now engage in a parallel transition, for example expressing the writing of shared data.
$$
\begin{array}{c}
  \Tpar {\Trinst {i.r} c \Xs {\TrteIrvWrite s v k}}
        {\Tirv {i.s} \_} \vspace{0.6ex} \\
  \red {\say {i.s} {\Tirvw v}} \\
  \Tpar {\Trinst {i.r} c \Xs {\Tap k \Tvoid}}
        {\Tirv {i.s} v}
\end{array}
$$
The corresponding read transition looks as follows:
$$
\begin{array}{c}
  \Tpar {\Trinst {i.r} c \Xs {\TrteIrvRead s k}}
        {\Tirv {i.s} v} \vspace{0.6ex} \\
  \red {\say {i.s} {\Tirvr v}} \\
  \Tpar {\Trinst {i.r} c \Xs {\Tap k v}}
        {\Tirv {i.s} v}
\end{array}
$$
Section~\ref{sect:Ignore} will show how examples like these can be extended to the case when both participating processes are embedded in arbitrarily large parallel process compositions.

@@

\section{Acknowledgements}

This work was mainly carried out in the project ``Resource Aware
Functional Programming (RAWFP)'' funded by the Swedish Foundation for
Strategic Research.
